Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-02-04 manuscript/climate_sensitivity_draft.Rmd
working on LS comments
#65

Crown positions ($CP$) were recorded for all cored trees that remained standing during the growing season of 2018. We followed the crown position protocol of @jennings_assessing_1999, whereby positions were ranked as dominant, codominant, intermediate, or suppressed. While some trees undoubtedly changed position in the 52 years between the 1966 drought and our observations in 2018, in this case the bias would be unlikely to result in false acceptance of our hypothesis (i.e., type I error unlikely; type II error possible). An analysis of crown position relative to height (Fig. 2d) and height changes since the beginning of the study period indicated that changes were fairly small relative to differences among canopy positions, although dominant and co-dominant trees were similar in height (Fig. S3). Specifically, average tree height growth was confined to ~0.82m from 1966 to 1977, ~1.45m from 1977 to 1999, and ~1.97m from 1999 to 2018.

@@

Models were run for all drought years combined (with year as a fixed effect) and for each drought year independently. In order to determine the relative importance of the traits alone (H2.1), we first tested the predictor variables independently against both height and $Rt$ given height's substantial influence. Variables were considered useful for predicting $Rt$ when AICc was reduced by $\ge$ 2 relative to the appropriate null model lacking that variable (Table 4).

@@

*Focal droughts and their characteristics*

@@

Community-level tree growth responses to these droughts were modest, with mean resistance values of 0.86, 0.84, and 0.86 for 1966, 1977, and 1999, respectively (Fig. 1b). All droughts yielded about 5% more trees experiencing negative growth ($Rt$ $\le$ 0.7) as defined by our drought threshold, specifically, 29% in 1966, 32% in 1977, and 27% in 1999. However, some trees exhibited increased growth: ($Rt$ > 1.0): 26% in 1966, 22% in 1977, and 26% in 1999.

@@

Overall, our analysis confirmed the tendency for larger-diameter trees to show greater reductions in growth during drought [@bennett_larger_2015], although there was no significant effect during 1977 or 1999 individually (Tables 1, 4). The same held true for $ln[H]$ as a univariate predictor (*H1a*; Tables 1, 4). When combined with other predictor variariables in our multivariate models, the top models usually included an effect of $ln[H]$, and its coefficient was consistently negative, as predicted (Tables 1, 5). We note that a non-signficant positive correlation between $ln[H]$ and $Rt$ for 1999 became negative in the context of the multivariate models, again supporting *H1a* (Table 1).

Crown position was generally correlated with $H$, but with substantial variation (Fig. 2d). Crown position was a much poorer predictor of $Rt$ than was $H$ (Table 4), lending little overall support to *H1b* (Table 1). When considered alone, $CP$ had a significant response only in the 1966 drought, during which trees with dominant $CP$ had the lowest $Rt$. When considered in conjunction with $H$, $CP$ resulted as a significant predictor only for the 1999 drought, during which suppressed and then intermediate trees had the lowest $Rt$. Crown position was included in **almost half of the top models, with mixed results as to how $Rt$ varied with $CP$ (Table 5). Most commonly in these multivariate models, as in the univariate models (Table 4), the resistance of dominant trees was less than that of co-dominant trees but higher that of suppressed trees**. Although $CP$ was at times a useful predictor of $Rt$, it overall had a weak effect relative to that of $H$.

@@

Our results partially support Hypothesis *H2.1*, where species' hydraulic traits --$XP$, $PLA_{dry}$, and $\pi_{tlp}$--were sometimes useful in explaining variation in drought responses, whereas $LMA$ and $WD$ never significantly associated with $Rt$ in the univariate models (all dAIC $\le$ 0.22; Tables 1,4,5). Therefore, these were excluded as candidate variables for the full multivariate models. In contrast, $XP$, $PLA_{dry}$, and $\pi_{tlp}$ all explained modest amounts of variation (dAIC > 1.0) in at least one drought (Table 4). $PLA_{dry}$ was a strong predictor for the 1966 drought and all droughts combined, with consistently negative coefficients. $\pi_{tlp}$ was not significant in the univariate models but had a consistently negative coefficient (Table 4). Whereas ring-porous species had highest $Rt$ overall and in the 1966 and 1999 droughts, they had lower $Rt$ in 1977. Results were similar in the context of multivariate models (Table 5). $PLA_{dry}$, and $\pi_{tlp}$ were each included in the top overall model and in the top model for two of the three individual droughts (Table 5).

>>>>

Crown position ($CP$)--a categorical variable including dominant, co-dominant, intermediate, and suppressed--was recorded for all cored trees that remained standing during the growing season of 2018 following the protocol of @jennings_assessing_1999. While some trees undoubtedly changed position in the 52 years between the 1966 drought and our observations in 2018, in this case the bias would be unlikely to result in false acceptance of our hypothesis (i.e., type I error unlikely; type II error possible). An analysis of crown position relative to height (Fig. 2d) and height changes since the beginning of the study period indicated that changes were fairly small relative to differences among canopy positions, although dominant and co-dominant trees were similar in height (Fig. S3). Specifically, average tree height growth was confined to ~0.82m from 1966 to 1977, ~1.45m from 1977 to 1999, and ~1.97m from 1999 to 2018.

@@

Models were run for all drought years combined (with year as a fixed effect) and for each drought year independently. In order to determine the relative importance of the traits alone (H2.1), we first tested the predictor variables independently against both height and $Rt$ given height's substantial influence. Variables were considered useful for predicting $Rt$ when AICc was reduced by $\ge$ 2 relative to the corresponding null model lacking that variable (Table 4).

@@

*Focal droughts and their characteristics* {**Lawren thinks most of this (1st 3 par) belong in the methods, that par 3 might not be really useful**}

@@

Community-level tree growth responses to these droughts were modest, with mean resistance values of 0.86, 0.84, and 0.86 for 1966, 1977, and 1999, respectively (Fig. 1b). All droughts yielded about 5% more trees experiencing negative growth ($Rt$ $\le$ 0.7) as defined by our drought threshold, specifically, 29% in 1966, 32% in 1977, and 27% in 1999. However, some individuals exhibited increased growth: ($Rt$ > 1.0): 26% in 1966, 22% in 1977, and 26% in 1999.

@@

Larger-diameter trees to showed greater reductions in growth during drought, although there was no significant effect during 1977 or 1999 individually (Tables 1, 4). The same held true for $ln[H]$ as a univariate predictor (*H1a*; Tables 1, 4). When combined with other predictor variables in our multivariate models, the top models usually included an effect of $ln[H]$, with negative coefficient (Tables 1, 5).

Crown position varied as expected with $H$ (dominant > co-dominant > intermediate > supressed), but with substantial variation (Fig. 2d). Crown position was a much poorer predictor of $Rt$ than was $H$ (Table 4), lending little overall support to *H1b* (Table 1). When considered alone, $CP$ had a significant response only in the 1966 drought, during which trees with dominant $CP$ had the lowest $Rt$. When considered in conjunction with $H$, $CP$ resulted as a significant predictor only for the 1999 drought, during which suppressed and then intermediate trees had the lowest $Rt$. Crown position was included in **almost half of the top models, with mixed results as to how $Rt$ varied with $CP$ (Table 5). Most commonly in these multivariate models, as in the univariate models (Table 4), the resistance of dominant trees was less than that of co-dominant trees but higher that of suppressed trees**. Although $CP$ was at times a useful predictor of $Rt$, it overall had a weak effect relative to that of $H$.

@@

Hydraulic traits, including $XP$, $PLA_{dry}$, and $\pi_{tlp}$, explained variation in drought responses, whereas $LMA$ and $WD$ never significantly associated with $Rt$ in the univariate models (Tables 1,4,5). $XP$, $PLA_{dry}$, and $\pi_{tlp}$ all explained modest amounts of variation (dAIC > 1.0) in at least one drought (Table 4). $PLA_{dry}$ was a strong predictor for the 1966 drought and all droughts combined, with consistently negative coefficients. $\pi_{tlp}$ had no significant effect in the univariate models, but coefficients were consistently negative (Table 4). Whereas ring-porous species had highest $Rt$ overall and in the 1966 and 1999 droughts, they had lower $Rt$ in 1977. Results were similar in the context of multivariate models (Table 5), and $PLA_{dry}$, and $\pi_{tlp}$ were each included in the top overall model and in the top model for two of the three individual droughts (Table 5).
